---
title: "Heart Attack Analysis & Prediction"
author: "Benjamin Barnett & Krishna Piratla"
date: "5/4/2021"
output: html_document
---

```{r setup, include=FALSE}
library(MASS)
library(ROCR)
library(class)
library(gam)
library(randomForest)
library(tidyverse)
library(ggpubr)
```


After reading and cleaning the data, we examine plots for each feature.

```{r, echo=FALSE}
heart <- rheart <- read.csv("C:/Users/Barnett/Downloads/heart.csv")
colnames(heart)[10] <- "old.peak"

index_remove <- NULL
for(i in seq(nrow(heart))){
if(heart$thall[i] == 0 | heart$caa[i] == 4){
	index_remove <- c(index_remove, i)
	}
if(heart$restecg[i] == 2){
	heart$restecg[i] <- 1
	}
}

heart <- heart[-index_remove,]

heart$sex <- factor(heart$sex)
levels(heart$sex) <- c("_female", "_male")

heart$cp <- factor(heart$cp)
levels(heart$cp) <- c("_typical","_atypical","_non-anginal","_asymptomatic")

heart$fbs <- factor(heart$fbs)

heart$restecg <- factor(heart$restecg)
levels(heart$restecg) <- c("_normal","_stt")

heart$exng <- factor(heart$exng)

heart$slp <- factor(heart$slp)
levels(heart$slp) <- c("_up","_flat","_down")

heart$thall <- factor(heart$thall)
levels(heart$thall) <- c("_normal","_fixed","_reversable")

heart$output <- factor(heart$output)
```


```{r, echo=FALSE}
ggplot(heart, aes(age, fill=output)) + geom_histogram(binwidth=2) + labs(fill="Disease", x="Age", y="Number of Patients") + scale_fill_manual(values=c("#11D164","#F8766D"))
```


```{r, echo=FALSE}
a <- ggplot(heart, aes(trtbps, fill=output)) + geom_histogram(binwidth=5) + labs(fill="Disease", x="Resting Blood Pressure", y="Number of Patients") + scale_fill_manual(values=c("#11D164","#F8766D"))
b <- ggplot(heart, aes(chol, fill=output)) + geom_histogram(binwidth=20) + labs(fill="Disease", x="Serum Cholesterol in mg/dl", y="Number of Patients") + scale_fill_manual(values=c("#11D164","#F8766D"))
c <- ggplot(heart, aes(thalachh, fill=output)) + geom_histogram(binwidth=5) + labs(fill="Disease", x="Maximum Heart Rate Achieved", y="Number of Patients") + scale_fill_manual(values=c("#11D164","#F8766D"))
d <- ggplot(heart, aes(old.peak, fill=output)) + geom_histogram(binwidth=.5) + labs(fill="Disease", x="ST Depression Induced by Exercise", y="Number of Patients") + scale_fill_manual(values=c("#11D164","#F8766D"))

ggarrange(a, b, c, d, ncol=2, nrow=2)
```


```{r, echo=FALSE}
e <- ggplot(heart, aes(sex, fill=output)) + geom_bar() + labs(fill="Disease", x="Sex", y="Number of Patients") + scale_x_discrete(labels=c("Female","Male")) + scale_fill_manual(values=c("#11D164","#F8766D"))
f <- ggplot(heart, aes(cp, fill=output)) + geom_bar() + labs(fill="Disease", x="Chest Pain", y="Number of Patients") + scale_x_discrete(labels=c("Typ", "Atyp", "Non-Ang", "Asymp")) + scale_fill_manual(values=c("#11D164","#F8766D"))
g <- ggplot(heart, aes(fbs, fill=output)) + geom_bar() + labs(fill="Disease", x="Fasting Blood Sugar > 120 mg/dl", y="Number of Patients") + scale_x_discrete(labels=c("False","True")) + scale_fill_manual(values=c("#11D164","#F8766D"))
h <- ggplot(heart, aes(restecg, fill=output)) + geom_bar() + labs(fill="Disease", x="Electrocardiogram Results", y="Number of Patients") + scale_x_discrete(labels=c("Normal", "ST-T")) + scale_fill_manual(values=c("#11D164","#F8766D"))

ggarrange(e,f,g,h, ncol=2, nrow=2)
```


```{r, echo=FALSE}
i <- ggplot(heart, aes(exng, fill=output)) + geom_bar() + labs(fill="Disease", x="Exercise Induced Angina", y="Number of Patients") + scale_x_discrete(labels=c("No", "Yes")) + scale_fill_manual(values=c("#11D164","#F8766D"))
j <- ggplot(heart, aes(slp, fill=output)) + geom_bar() + labs(fill="Disease", x="Peak Exercise ST Segment Slope",y="Number of Patients") + scale_x_discrete(labels=c("Upsloping","Flat","Downsloping")) + scale_fill_manual(values=c("#11D164","#F8766D"))
k <- ggplot(heart, aes(caa, fill=output)) + geom_bar() + labs(fill="Disease", x="Number of Major Vessels", y="Number of Patients") + scale_fill_manual(values=c("#11D164","#F8766D"))
l <- ggplot(heart, aes(thall, fill=output)) + geom_bar() + labs(fill="Disease", x="Thallium Stress Results", y="Number of Patients") + scale_x_discrete(labels=c("Normal", "Fixed", "Reversible")) + scale_fill_manual(values=c("#11D164","#F8766D"))

ggarrange(i,j,k,l, ncol=2, nrow=2)
```


The response (Coronary Artery Narrowing > 50%) is balanced, as 160 patients have heart disease and 136 patients do not have heart disease.

```{r, echo=FALSE}
table(heart$output)
```


Fitting logistic model with 50% training data.

```{r, echo=FALSE}
set.seed(1)
train <- sample(1:nrow(heart), 0.5*nrow(heart))
output.test <- heart[-train,]$output

glm.fit <- glm(output ~., heart[train,], family=binomial)
glm.prb <- predict(glm.fit, heart[-train,], type="response")
	
roc.prd <- prediction(glm.prb, output.test)
roc.prf <- performance(roc.prd,"tpr","fpr")

plot(roc.prf, colorize=TRUE)
cat("AUC:", as.numeric((performance(roc.prd,"auc"))@y.values), "\n")
```


```{r}
glm.prd <- rep("0", 148); glm.prd[glm.prb > .2] <- "1"
table(glm.prd, output.test)
```


```{r, echo=FALSE}
cat("Classification Accuracy:", (table(glm.prd, output.test)[1] + table(glm.prd, output.test)[4])/148, "\n")
```


```{r}
glm.prd <- rep("0", 148); glm.prd[glm.prb > .5] <- "1"
table(glm.prd, output.test)
```


```{r, echo=FALSE}
cat("Classification Accuracy:", (table(glm.prd, output.test)[1] + table(glm.prd, output.test)[4])/148, "\n")
```


```{r}
glm.prd <- rep("0", 148); glm.prd[glm.prb > .8] <- "1"
table(glm.prd, output.test)
```


```{r, echo=FALSE}
cat("Classification Accuracy:", (table(glm.prd, output.test)[1] + table(glm.prd, output.test)[4])/148, "\n")
```


```{r}
summary(glm.fit)
```


Fitting LDA model with 50% training data.

```{r, echo=FALSE}
set.seed(1)
train <- sample(1:nrow(heart), 0.5*nrow(heart))
output.test <- heart[-train,]$output

lda.fit <- lda(output ~., heart[train,])
lda.prb <- predict(lda.fit, heart[-train,], type="response")
	
roc.prd <- prediction(lda.prb$posterior[,2], output.test)
roc.prf <- performance(roc.prd,"tpr","fpr")

plot(roc.prf, colorize=TRUE)
cat("AUC:", as.numeric((performance(roc.prd,"auc"))@y.values), "\n")
```


```{r}
lda.prd <- rep("0", 148); lda.prd[lda.prb$posterior[,2] > .2] <- "1"
table(lda.prd, output.test)
```


```{r, echo=FALSE}
cat("Classification Accuracy:", (table(lda.prd, output.test)[1] + table(lda.prd, output.test)[4])/148, "\n")
```


```{r}
lda.prd <- rep("0", 148); lda.prd[lda.prb$posterior[,2] > .5] <- "1"
table(lda.prd, output.test)
```


```{r, echo=FALSE}
cat("Classification Accuracy:", (table(lda.prd, output.test)[1] + table(lda.prd, output.test)[4])/148, "\n")
```


```{r}
lda.prd <- rep("0", 148); lda.prd[lda.prb$posterior[,2] > .8] <- "1"
table(lda.prd, output.test)
```


```{r, echo=FALSE}
cat("Classification Accuracy:", (table(lda.prd, output.test)[1] + table(lda.prd, output.test)[4])/148, "\n")
```


Fitting QDA model with 50% training data.

```{r, echo=FALSE}
set.seed(1)
train <- sample(1:nrow(heart), 0.5*nrow(heart))
output.test <- heart[-train,]$output

qda.fit <- qda(output ~., heart[train,])
qda.prb <- predict(qda.fit, heart[-train,], type="response")
	
roc.prd <- prediction(qda.prb$posterior[,2], output.test)
roc.prf <- performance(roc.prd,"tpr","fpr")

plot(roc.prf, colorize=TRUE)
cat("AUC:", as.numeric((performance(roc.prd,"auc"))@y.values), "\n")
```


```{r}
qda.prd <- rep("0", 148); qda.prd[qda.prb$posterior[,2] > .2] <- "1"
table(qda.prd, output.test)
```


```{r, echo=FALSE}
cat("Classification Accuracy:", (table(qda.prd, output.test)[1] + table(qda.prd, output.test)[4])/148, "\n")
```


```{r}
qda.prd <- rep("0", 148); qda.prd[qda.prb$posterior[,2] > .5] <- "1"
table(qda.prd, output.test)
```


```{r, echo=FALSE}
cat("Classification Accuracy:", (table(qda.prd, output.test)[1] + table(qda.prd, output.test)[4])/148, "\n")
```


```{r}
qda.prd <- rep("0", 148); qda.prd[qda.prb$posterior[,2] > .8] <- "1"
table(qda.prd, output.test)
```


```{r, echo=FALSE}
cat("Classification Accuracy:", (table(qda.prd, output.test)[1] + table(qda.prd, output.test)[4])/148, "\n")
```


Fitting logistic GAM model on 50% training data.

```{r, echo=FALSE}
set.seed(1)
train <- sample(1:nrow(heart), 0.5*nrow(heart))
output.test <- heart[-train,]$output

gam.fit <- gam(output ~., data=heart[train,], family=binomial)
gam.prb <- predict(gam.fit, heart[-train,], type="response")

roc.prd <- prediction(gam.prb, output.test)
roc.prf <- performance(roc.prd,"tpr","fpr")

plot(roc.prf, colorize=TRUE)
cat("AUC:", as.numeric((performance(roc.prd,"auc"))@y.values), "\n")
```


```{r}
gam.prd <- rep("0", 148); gam.prd[gam.prb > .2] <- "1"
table(gam.prd, output.test)
```


```{r, echo=FALSE}
cat("Classification Accuracy:", (table(gam.prd, output.test)[1] + table(gam.prd, output.test)[4])/148, "\n")
```


```{r}
gam.prd <- rep("0", 148); gam.prd[gam.prb > .5] <- "1"
table(gam.prd, output.test)
```


```{r, echo=FALSE}
cat("Classification Accuracy:", (table(gam.prd, output.test)[1] + table(gam.prd, output.test)[4])/148, "\n")
```


```{r}
gam.prd <- rep("0", 148); gam.prd[gam.prb > .8] <- "1"
table(gam.prd, output.test)
```


```{r, echo=FALSE}
cat("Classification Accuracy:", (table(gam.prd, output.test)[1] + table(gam.prd, output.test)[4])/148, "\n")
```


```{r}
summary(gam.fit)
```


Fitting random forest with 50% training data.

```{r, echo=FALSE}
set.seed(1)
train <- sample(1:nrow(heart), 0.5*nrow(heart))
output.test <- heart[-train,]$output

rnf.fit <- randomForest(output ~., data=heart[train,])
rnf.prb <- predict(rnf.fit, heart[-train,], type="prob")

roc.prd <- prediction(rnf.prb[,2], output.test)
roc.prf <- performance(roc.prd,"tpr","fpr")

plot(roc.prf, colorize=TRUE)
cat("AUC:", as.numeric((performance(roc.prd,"auc"))@y.values), "\n")
```


```{r}
rnf.prd <- rep("0", 148); rnf.prd[rnf.prb[,2] > .2] <- "1"
table(rnf.prd, output.test)
```


```{r, echo=FALSE}
cat("Classification Accuracy:", (table(rnf.prd, output.test)[1] + table(rnf.prd, output.test)[4])/148, "\n")
```


```{r}
rnf.prd <- rep("0", 148); rnf.prd[rnf.prb[,2] > .5] <- "1"
table(rnf.prd, output.test)
```


```{r, echo=FALSE}
cat("Classification Accuracy:", (table(rnf.prd, output.test)[1] + table(rnf.prd, output.test)[4])/148, "\n")
```


```{r}
rnf.prd <- rep("0", 148); rnf.prd[rnf.prb[,2] > .8] <- "1"
table(rnf.prd, output.test)
```


```{r, echo=FALSE}
cat("Classification Accuracy:", (table(rnf.prd, output.test)[1] + table(rnf.prd, output.test)[4])/148, "\n")
```


```{r}
varImpPlot(rnf.fit, main="")